% The research presented in this work has revealed several key differences between models for a spring-operated and pilot-operated PRVs. However, there are significant areas revealed by these models which need greater exploration in further work.

% Firstly, the quarter-wave model presented which shows many of the unstable behaviours of a pilot-operated PRV must be validated as a model. Significantly more computational fluid dynamics (CFD) models and experimental results are needed to aid in the modelling process of the PRV. One key finding of this report is that the main piston closes because the valve-open equilibrium is unstable (\cref{sec: Prog}).

\newpage
\begin{itemize}
    \item More CFD and experimental results required:
    \begin{itemize}
        \item Validate the quarter-wave model
        \item Is the valve open equilibrium unstable? Can the valve open further?
        \item Identify flutter and chatter behaviour for given parameters
    \end{itemize}
    \item Further exploration of this analytical model
    \begin{itemize}
        \item Why are the oscillations not at the quarter-wave frequency?
        \item Does the stability boundary change by allowing the tank pressure to vary?
    \end{itemize}
    \item Extension this analytical model
    \begin{itemize}
        \item Add pilot valve dynamics - allows 
        \item Immediately after pilot valve closes, is the initial conditions below the unstable equilibrium?
        \item Does the model present other instabilities such as cycling?
        \item How does the pilot valve effect the flutter and chatter instability?
    \end{itemize}
    \item Experimental results
        \begin{itemize}
        \item Validate the quarter-wave model
        \item Is the valve open equilibrium unstable? Can the valve open further?
        \item Offer opportunity to compare calculated stability boundaries against "true" data
    \end{itemize}
\end{itemize}