\chapter{Introduction}

Pressure relief valves (PRVs) are an important safety feature in many pressurised systems, forming a final defence against over pressurisation. A fluid operating at a higher pressure than designed can lead to decreases in efficiency for some processes. More severely, a fluid within a vessel reaching high pressures can cause an explosive release of fluid. This has two serious impacts. Firstly, severe damage can cause costly repairs, possibly including replacing the entire pressure tank. Not only the capital cost of repairs must be considered, but also the potential loss of revenue for being unable to use a process. Secondly, it could prove fatal to an unfortunate operator or even the nearby public. The Piper Alpha disaster was caused when a PRV was removed for maintenance, leading to the death of 167 people~\cite{Drysdale1998TheStudy}.

The simplest type of PRV is the direct spring-operated PRV, shown in \cref{fig: IntroSpring}. These are widely used, and rely on a spring to hold a poppet in place. As the pressure increases, a greater force acts upon the poppet to open the valve. Hence, the poppet will rise which allows fluid to escape causing the pressure in the tank to be relieved. One advantage of PRVs is they prevent an unnecessary waste of fluid because the valve will close once a safe pressure is reached. However, spring-operated PRVs do not scale well for larger vessels as larger flow rates are needed requiring a very large physical space.

\input{Sections/Stuff/IntroDiagram.tex}

% \vspace{-10pt}
A more complex design of PRV is the pilot-operated PRV, shown in \cref{fig: IntroPilot}. This allows fluid both above and below the piston, and so uses the fluid pressure to keep the valve shut. When the pressure becomes too great, a small pilot valve opens which reduces the pressure above the piston. Hence, the main valve opens allowing the fluid to escape and relieve the pressure in the vessel. This has several advantages. Even for large tanks, no physical spring is needed, so the valve can be physically smaller than a spring-operated PRV.

Unfortunately, both spring and pilot operated PRVs do not always behave as expected. Several forms of instability exist, ranging from a desirable blow-down effect to the most damaging chatter instability~\cite{Hos2017DynamicRecommendations}. A blow-down effect is when the PRV closes at a lower pressure than that at which it opens, called the set pressure. This is desirable as it can help the valve remain slightly open rather than cycling, where the valve intermittently opens and closes, typically at frequencies less than $1 \, \si{Hz}$.

The more damaging chatter instability is characterised by high frequency oscillations, typically $\gg 1 \, \si{Hz}$, in which the poppet or piston undergoes impacts. These oscillations contain large amounts of energy, which can cause severe damage to the PRV. This damage can decrease the performance of the valve, or even cause it to fail completely. These consequences have serious repercussions for the pressurised vessel as the final safety precaution is compromised, meaning the dangerous effects of over pressurisation may occur.

% This project aims to study the various types of instability which can occur for a pilot-operated PRV. A particular focus is placed on identifying the damaging chatter instability. If successful in establishing a stability criteria, this work could allow a pilot-operated PRV to be designed to avoid chatter. However, first a review of the current understanding of instability in both spring-operated and pilot-operated PRVs will be presented.

\input{Sections/literature_review.tex}

\section{Project aims and objectives}

As the review of the current literature demonstrates, very little research has been performed for pilot-operated PRVs. Hence, this project aims to study the various types of instability which can occur for a pilot-operated PRV. A particular focus is placed on identifying the damaging chatter instability. To study the instabilities of a pilot-operated PRV, first a suitable model must be derived which includes the fluid dynamics in the inlet piping. Then, some of the methods used to understand flutter and chatter in the simpler direct spring-operated PRV will be applied to this model. If successful in establishing a stability criteria, this work could allow a pilot-operated PRV to be designed to avoid chatter.

\section{Plan of report}

The rest of this report is outlined as follows. \Cref{sec: Prog} will detail the derivation and analysis of a simple model describing the closing of a pilot-operated PRV, ignoring the inlet piping. It is then necessary to consider the effect of the inlet piping, so \cref{sec: QWM} will present the derivation of a quarter-wave model. This model will then be used to calculate an analytical stability boundary for the chatter instability. \Cref{sec: Continuation} will use numerical continuation techniques to analysis the quarter-wave model. Next, \cref{sec: QWSim} includes numerical simulations of the quarter-wave model which demonstrate flutter, chatter and stable operation. Finally, this report concludes by explaining the possibilities of further work and will draw any final conclusions.

%what are you working on?
%what is the real-world challenge?
%why should anyone care?